---
title: "Examples --- Canonical Form of LP"
---

```{r message = FALSE, warning = FALSE, includet = FALSE}

library(simplexR)
library(gMOIP) # for visualisations

```

## Standard Case

--- Werners, Brigitte - Grundlagen des Operations Research (p. 60)

$$
\begin{array}{rrcrcr}
\max  & 3x_1 & + & 2x_2 & \\
s.t.  & 2x_1 & + & 1x_2 & \leq & 22 \\
      & 1x_1 & + & 2x_2 & \leq & 23 \\
      & 4x_1 & + & 1x_2 & \leq & 40 \\
      &  x_1,&  &  x_2 & \geq & 0
\end{array}
$$

```{r}

# standard case 
# optimal solution: x = (7, 8), z = 37

A <- matrix(c(
  2, 1,
  1, 2,
  4, 1
), nrow = 3, ncol = 2, byrow = TRUE, dimnames = list(c("R1", "R2", "R3"), c("x1", "x2")))

b <- c(22, 23, 40)
c <- c(3, 2)
sense <- 1

simplexR(A, b, c, sense)

```

```{r}

# visualisation
gMOIP::plotPolytope(A = A, b = b, obj = c, crit = "max", plotOptimum = TRUE, labels = "coord")

```


## Standard Case---Exponential Time Complexity 

$$
\begin{array}{rrcrcr}
\max  & 2x_1 & + & 1x_2 & \\
s.t.  & 1x_1 &   &      & \leq &  5 \\
      & 4x_1 & + & 1x_2 & \leq & 25 \\
      &  x_1,&   &  x_2 & \geq & 0
\end{array}
$$

```{r}

# standard case
# optimal solution: x = (0, 25), z = 25

A <- matrix(c(
  1, 0,
  4, 1
), nrow = 2, ncol = 2, byrow = TRUE, dimnames = list(c("R1", "R2"), c("x1", "x2")))

b <- c(5, 25)
c <- c(2, 1)

simplexR(A, b, c)

```

```{r}

# visualisation
gMOIP::plotPolytope(A = A, b = b, obj = c, crit = "max", plotOptimum = TRUE, labels = "coord")

```

## Unbounded Feasible Region, Optimal Solution

$$
\begin{array}{rrcrcr}
\max  &-2x_1 & + & 3x_2 & \\
s.t.  &-1x_1 & + & 1x_2 & \leq & 2 \\
      &-1x_1 & + & 2x_2 & \leq & 6 \\
      &      &   & 1x_2 & \leq & 5 \\
      &  x_1,&  &  x_2 & \geq & 0
\end{array}
$$

```{r}

# unbounded feasible region, optimal solution exists
# optimal solution: x = (2, 4), z = 8

A <- matrix(c(
  -1, 1,
  -1, 2,
  0, 1
), nrow = 3, ncol = 2, byrow = TRUE, dimnames = list(c("R1", "R2", "R3"), c("x1", "x2")))

b <- c(2, 6, 5)
c <- c(-2, 3)
sense <- 1

simplexR(A, b, c, sense)

```

```{r}

# visualisation is a bit wrong!
gMOIP::plotPolytope(A = A, b = b, obj = c, crit = "max", plotOptimum = TRUE, labels = "coord")

```

## Unbounded Feasible Region, no Optimal Solution 

--- Werners, Brigitte - Grundlagen des Operations Research (p. 60)

$$
\begin{array}{rrcrcr}
\max  & 2x_1 & + & 4x_2 & \\
s.t.  &-2x_1 & + & 3x_2 & \leq & 12 \\
      &-1x_1 & + & 3x_2 & \leq & 18 \\
      &  x_1,&  &  x_2 & \geq & 0
\end{array}
$$

```{r}

# unbounded feasible region, no optimal solution exists

A <- matrix(c(
  -2, 3,
  -1, 3
), nrow = 2, ncol = 2, byrow = TRUE)

b <- c(12, 18)
c <- c(2, 4)
sense <- 1

simplexR(A, b, c, sense)

```

## Dual Degeneracy 

--- Werners, Brigitte - Grundlagen des Operations Research (p. 62)

$$
\begin{array}{rrcrcr}
\max  & 3x_1 & + &1.5x_2 & \\
s.t.  & 2x_1 & + & 1x_2 & \leq & 22 \\
      & 1x_1 & + & 2x_2 & \leq & 23 \\
      & 4x_1 & + & 1x_2 & \leq & 40 \\
      &  x_1,&  &  x_2 & \geq & 0
\end{array}
$$

```{r}

# dual degeneracy
# optimal solution: x = (9, 4), z = 33
# optimal solution: x = (7, 8), z = 33

A <- matrix(c(
  2, 1,
  1, 2,
  4, 1
), nrow = 3, ncol = 2, byrow = TRUE)

b <- c(22, 23, 40)
c <- c(3, 1.5)
sense <- 1

simplexR(A, b, c, sense)

```

```{r}

# visualisation
gMOIP::plotPolytope(A = A, b = b, obj = c, crit = "max", plotOptimum = TRUE, labels = "coord")

```

## Primal Degeneracy

--- Werners, Brigitte - Grundlagen des Operations Research (p. 65)

$$
\begin{array}{rrcrcr}
\max  & 3x_1 & + & 2x_2 & \\
s.t.  & 2x_1 & + & 1x_2 & \leq & 22 \\
      & 1x_1 & + & 2x_2 & \leq & 23 \\
      & 4x_1 & + & 1x_2 & \leq & 40 \\
      & 2x_1 & + & 0.75x_2 & \leq & 21 \\
      &  x_1,&  &  x_2 & \geq & 0
\end{array}
$$

```{r}

# primal degeneracy
# optimal solution: x = (7, 8), z = 37

A <- matrix(c(
  2, 1,
  1, 2,
  4, 1,
  2, 3 / 4
), nrow = 4, ncol = 2, byrow = TRUE)

b <- c(22, 23, 40, 21)
c <- c(3, 2)
sense <- 1

simplexR(A, b, c, sense)

```

```{r}

# visualisation
gMOIP::plotPolytope(A = A, b = b, obj = c, crit = "max", plotOptimum = TRUE, labels = "coord")

```
